Investigating magnetic dynamo evolution with TESS field dwarfs

### 2017-10-04 ms.tex

\documentclass[useAMS, usenatbib, preprint, 12pt]{aastex}
\usepackage{cite, natbib}
\usepackage{float}
\usepackage{epsfig}
\usepackage{cases}
\usepackage[section]{placeins}
\usepackage{graphicx, subfigure}
\usepackage{color}
\usepackage{bm}

@@

\section{Introduction}

% Age is a fundamental stellar parameter of great interest to at galactic
% archaeologists and exoplaneteers alike but unfortunately it is a difficult
% attribute to measure for main sequence F, G, K, and M stars in the field.
Of the measurable properties for a large ensemble of field stars, rotation
periods contain the most information about stellar age, and provide the best
leverage for advancing our knowledge of galactic archeology as well as
exoplanet population demographics via gyrochronology.
Angular momentum is carried away though magnetically driven stellar winds,
which slows the star’s rotation over cosmic time.
This rotation-based `clock' is known as gyrochronology.
Cool spots on the star’s surface rotate in-to and out-of view, creating small
amplitude ($\pm$ 1\%) quasi-periodic changes in the stellar brightness.
Using precise light curves available from the \TESS\ mission, we expect to
measure the rotation periods of around ... stars.
For many of these stars it will be possible to infer an age from their
rotation periods via gyrochronology.

@@

\section{Scientific Justification}

@@

\begin{figure}
\begin{center}
\includegraphics[width=3in, clip=true]{Davenport.png}
\caption{From Davenport (2017).
    Rotation period as a function of effective
temperature for the full McQuillan et al., (2014) Kepler sample in black, and
the subset of these stars that also feature in the TGAS Gaia DR1 catalogue in
blue. Contaminating giants have been removed from the blue sample and the
rotation period bimodality is revealed to exist across all temperatures shown.
    The red line is a 600 Myr rotational isochrone (also known as a gyrochrone).}
\label{fig:davenport}
\end{center}
\end{figure}

@@

\section{Technical Feasibility}
The extraction of a rotation period from a light curve can be as simple as
computing a Lomb-Scargle periodogram or, in cases where the signal is less
clear, can be inferred via modeling the correlation between data points.
This latter approach could involve either computing an autocorrelation
function \citep{mcquillan2013}, or fitting a Gaussian process to the time
series \citep{angus2017, foreman-mackey2017}.
In cases where signals have long periods and low amplitudes, care is needed to
separate real astrophysical signal from instrumental systematics.
The measurement of rotation periods is less sensitive to crowding and source
confusion than exoplanet transit characterization because the rotation period
is not effected by photometric dilution.
We will apply two complementary methods to extract and calibrate light curves
from the TESS FFIs.
First, for bright or isolated targets, we will follow \citet{montet2017} to
estimate aperture shapes and perform aperture photometry for bright sources.
Using an ensemble of sources, we will de-trend these light curves using a
modified version of \textsf{everest} \citep{luger2016, luger2017} designed to
preserve photometric signatures of rotation.
This will be achieved by fitting for the systematic effects in the light curve
using the \textsf{everest} model simultaneously with a Gaussian Process model
for the astrophysical variation.
Both \citet{aigrain2016} and \citet{luger2016} demonstrated that this can
preserve stellar variability signals and we will use the \textsf{celerite}
algorithm \citep{dfm2017} to scale the computations to the size of TESS FFI
datasets.
The precision of existing photometric de-trending methods degrades in crowded
fields \citep[for example,][]{luger2017}.
However, to make robust measurements of rotation periods, we do not need
absolutely calibrated light curves.
Therefore, in crowded fields, we will apply a difference imaging method that
was developed for the K2 Campaign 9 microlensing project \citep{henderson2016}
based on the \textsf{CPM} \citep{wang2016} to robustly measure the photometric
variations of crowded sources.
Unlike standard difference imaging methods, this procedure does not require a
reference image.
Instead, a causal data-driven model is built to predict the time series in
every pixel taking systematic effects into account and the residuals between
the observations and the model predictions provide an estimate of the
astrophysical variability in each pixel.
We will tune this method preserve rotation signals and apply it to detect
rotation across the FFIs.

@@

\section{Expected Impact}
We will provide rotation periods for all two-minute cadence and FFI targets.

\section{Budget Justification}
PI Angus intends to use the budget to employ a student for 3-4 months.
In this time the student will assist in extracting light curves from the FFIs,
measuring rotation periods and building a rotation period catalog.
The student will also be involved in the scientific project of their choosing:
either the rotation period bimodality, gyrochronology of low-mass stars, or
gyrochronology of comoving pairs.

\bibliography{references}

>>>>

\documentclass[letterpaper,12pt,preprint]{hack_aastex}

\input{dfm_stylez}
\pagestyle{myheadings}
\markright{\textsf{\footnotesize %
    Investigating magnetic dynamo evolution with TESS field dwarfs}}

%\documentclass[useAMS, usenatbib, preprint, 12pt]{aastex}
%\usepackage{cite, natbib}
%\usepackage{float}
%\usepackage{epsfig}
%\usepackage{cases}
%\usepackage[section]{placeins}
%\usepackage{graphicx, subfigure}
%\usepackage{color}
%\usepackage{bm}

@@

%\paragraph{Introduction}

% Age is a fundamental stellar parameter of great interest to at galactic
% archaeologists and exoplaneteers alike but unfortunately it is a difficult
% attribute to measure for main sequence F, G, K, and M stars in the field.
Of the measurable properties for a large ensemble of field stars, rotation
periods contain the most information about stellar age, and provide the best
leverage for advancing our knowledge of galactic archeology as well as
exoplanet population demographics via gyrochronology.
Angular momentum is carried away though magnetically driven stellar winds,
which slows the star's rotation over cosmic time.
This rotation-based `clock' is known as gyrochronology.
Cool spots on the star's surface rotate in-to and out-of view, creating small
amplitude ($\pm$ 1\%) quasi-periodic changes in the stellar brightness.
Using precise light curves available from the \TESS\ mission, we expect to
measure the rotation periods of around ... stars.
For many of these stars it will be possible to infer an age from their
rotation periods via gyrochronology.

@@

\paragraph{Scientific Justification}

@@

\ssfigure{Davenport.png}{0.5}{%
From Davenport (2017).
    Rotation period as a function of effective
temperature for the full McQuillan et al., (2014) Kepler sample in black, and
the subset of these stars that also feature in the TGAS Gaia DR1 catalogue in
blue. Contaminating giants have been removed from the blue sample and the
rotation period bimodality is revealed to exist across all temperatures shown.
    The red line is a 600 Myr rotational isochrone (also known as a gyrochrone).
\label{fig:davenport}
}
%\begin{figure}
%\begin{center}
%\includegraphics[width=3in, clip=true]{Davenport.png}
%\caption{From Davenport (2017).
%    Rotation period as a function of effective
%temperature for the full McQuillan et al., (2014) Kepler sample in black, and
%the subset of these stars that also feature in the TGAS Gaia DR1 catalogue in
%blue. Contaminating giants have been removed from the blue sample and the
%rotation period bimodality is revealed to exist across all temperatures shown.
%    The red line is a 600 Myr rotational isochrone (also known as a gyrochrone).}
%\label{fig:davenport}
%\end{center}
%\end{figure}

@@

\paragraph{Technical Feasibility}
The extraction of a rotation period from a light curve can be as simple as
computing a Lomb-Scargle periodogram or, in cases where the signal is less
clear, can be inferred via modeling the correlation between data points.
This latter approach could involve either computing an autocorrelation
function \citep{mcquillan2013}, or fitting a Gaussian process to the time
series \citep{angus2017, foreman-mackey2017}.
In cases where signals have long periods and low amplitudes, care is needed to
separate real astrophysical signal from instrumental systematics.
The measurement of rotation periods is less sensitive to crowding and source
confusion than exoplanet transit characterization because the rotation period
is not effected by photometric dilution.
We will apply two complementary methods to extract and calibrate light curves
from the TESS FFIs.
First, for bright or isolated targets, we will follow \citet{montet2017} to
estimate aperture shapes and perform aperture photometry for bright sources.
Using an ensemble of sources, we will de-trend these light curves using a
modified version of \textsf{everest} \citep{luger2016, luger2017} designed to
preserve photometric signatures of rotation.
This will be achieved by fitting for the systematic effects in the light curve
using the \textsf{everest} model simultaneously with a Gaussian Process model
for the astrophysical variation.
Both \citet{aigrain2016} and \citet{luger2016} demonstrated that this can
preserve stellar variability signals and we will use the \textsf{celerite}
algorithm \citep{dfm2017} to scale the computations to the size of TESS FFI
datasets.
The precision of existing photometric de-trending methods degrades in crowded
fields \citep[for example,][]{luger2017}.
However, to make robust measurements of rotation periods, we do not need
absolutely calibrated light curves.
Therefore, in crowded fields, we will apply a difference imaging method that
was developed for the K2 Campaign 9 microlensing project \citep{henderson2016}
based on the \textsf{CPM} \citep{wang2016} to robustly measure the photometric
variations of crowded sources.
Unlike standard difference imaging methods, this procedure does not require a
reference image.
Instead, a causal data-driven model is built to predict the time series in
every pixel taking systematic effects into account and the residuals between
the observations and the model predictions provide an estimate of the
astrophysical variability in each pixel.
We will tune this method preserve rotation signals and apply it to detect
rotation across the FFIs.

@@

\paragraph{Expected Impact}
We will provide rotation periods for all two-minute cadence and FFI targets.

\paragraph{Budget Justification}
PI Angus intends to use the budget to employ a student for 3-4 months.
In this time the student will assist in extracting light curves from the FFIs,
measuring rotation periods and building a rotation period catalog.
The student will also be involved in the scientific project of their choosing:
either the rotation period bimodality, gyrochronology of low-mass stars, or
gyrochronology of comoving pairs.

\paragraph{References}

{\footnotesize
\bibliography{references}
}
